Model behavior analysis

### 2016-10-13 Rmd/analyze.Rmd
update

```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T}
# Import data from Marcelino et al. 2013
md <- read.csv("data/marcelino2013_fig2a.csv")
md$S.R <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
# Plot data
plot(md$A ~ md$S.R, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
     ylab="internal/external light", ylim=c(1,3.2),
     xlab="Symbiont density (CmolS/CmolR)",
     main="Light Amplification")
legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
# Fit exponential decay model and plot fitted values
mod <- nls(A ~ x + y * exp(z * S.R), data=md, start=list(x=1, y=2, z=-3))
fit <- predict(mod, list(S.R=seq(0, 0.5, 0.01)))
lines(fit ~ seq(0, 0.5, 0.01))
# Get parameters of fit
formula(mod); coef(mod)
```

@@

# Run to steady state and plot results
ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 100)), at=at, food=0, runtime=T)
plot_steady_states(ssdat)
```
With default parameters, steady state values of specific growth are higher than usually observed in corals (skeletal growth), but patterns across environments are consistent with literature.

### Steady states values with default parameters + feeding
```{r ss_feeding, cache=T, fig.width=5, fig.height=4.4}
# Set parameters
defpars <- def_pars()  # Get default parameters

@@

# Run to steady state and plot results
ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 100)), at=at, food=5e-6, runtime=T)
plot_steady_states(ssdat)
```
With default parameters, steady state values of specific growth are higher than usually observed in corals (skeletal growth), but patterns across environments are consistent with literature.

@@

# Run to steady state and plot results
ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS"), c(1, 200)), at=at, food=0, runtime=T)
plot_steady_states(ssdat)
```
With some feeding (more realistic), can maintain growth with very low symbionts (bleached) at very high or low light with high nutrients.

### Vary Light
```{r varyL, eval=F, include=F}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.2))#
plot_run(run)
```

### Vary nutrients
```{r varyN, eval=F, include=F}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(20,20,2), N=c(1e-7,4e-6,2), X=c(5e-6,5e-6,0))

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.25))#
plot_run(run)
```

### Vary nutrients @ very low light
```{r varyN_lowL, eval=F, include=F}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

@@

### Steadily increasing light
```{r incrL, eval=F, include=F}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.1))#
plot_run(run)
```

### Steadily increasing light, + high N
```{r incrL_highN, eval=F, include=F}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(15,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, "initS", 0.5))#
plot_run(run)
```

### Both on same plot
```{r incrL_multiline, eval=F, include=F}
time <- seq(1, 365, 0.5)  # Set time
defpars <- def_pars()  # Get default parameters

# Run 1
env1 <- init_env(time=time, L=c(15,60,0), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", 0.1))
# Run 2
env2 <- init_env(time=time, L=c(15,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))
run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", 0.5))
# Run 3
env3 <- init_env(time=time, L=c(15,60,0), N=c(4e-6,4e-6,2), X=c(5e-6,5e-6,0))
run3 <- run_coral(time=time, env=env3, pars=replace(defpars, "initS", 0.5))
# Run 4
env4 <- init_env(time=time, L=c(15,60,0), N=c(1e-7,1e-7,2), X=c(0,0,0))
run4 <- run_coral(time=time, env=env4, pars=replace(defpars, "initS", 0.2))

# Plot
plot(NA, xlim=range(time), ylim=c(0,0.5))
with(run1, lines(time, S$S/H$H, lty=1))
with(run2, lines(time, S$S/H$H, lty=2))
with(run3, lines(time, S$S/H$H, lty=3))
with(run4, lines(time, S$S/H$H, lty=4))
```

```{r incrL_multiline1, include=F, eval=F}
time <- seq(1, 200, 0.5)  # Set time
defpars <- def_pars()  # Get default parameters
defpars <- replace(defpars, "kNPQ", 112)

# Run 1
env1 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,2), X=c(5e-6,5e-6,0))
run1 <- run_coral(time=time, env=env1, pars=replace(defpars, "initS", 0.12))
# Run 2 - increase food
env2 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,2), X=c(10e-6,10e-6,0))
run2 <- run_coral(time=time, env=env2, pars=replace(defpars, "initS", 0.12))
# Run 3 - increase DIN
env3 <- init_env(time=time, L=c(15,40,2), N=c(2e-6,2e-6,2), X=c(10e-6,10e-6,0))
run3 <- run_coral(time=time, env=env3, pars=replace(defpars, "initS", 0.25))

# Plot
plot(time, env1$L, col=alpha("gold", 0.5), type="l", lwd=2, axes=F)
par(new=T)
plot(NA, xlim=range(time), ylim=c(0,0.3), yaxs="i")
with(run1, lines(time, S$S/H$H, lty=1))
with(run2, lines(time, S$S/H$H, lty=2))
with(run3, lines(time, S$S/H$H, lty=3))

@@

##### Plot run 1 symbiont PI curve
```{r run1_PI, cache=T, fig.height=3}
sym_PI(pars=defpars)
```

>>>>

![](../img/model_new.png)

### Light amplification & self-shading
```{r lightamp, fig.width=5, fig.height=5, eval=T, include=T}
# Import data from Marcelino et al. 2013
md <- read.csv("data/marcelino2013_fig2a.csv")
md$S.R <- md$density * cm2.m2 / CmolR.m2 * CmolS.cellS  # Convert units of symbiont density to CmolS/CmolR
md$A <- md$E + 1  # Convert "E"=excess light metric of Marcelino to "A"=amplification factor
# Plot data
plot(md$A ~ md$S.R, pch=c(21,23,22)[factor(md$spp)], bg=c("white", "black", "gray")[factor(md$spp)], 
     ylab="internal/external light", ylim=c(1,3.2),
     xlab="Symbiont density (CmolS/CmolR)",
     main="Light Amplification")
legend("topright", legend=levels(md$spp), pch=c(21,23,22), pt.bg=c("white", "black", "gray"), inset=0.02)
# Fit exponential decay model and plot fitted values
mod <- nls(A ~ x + y * exp(z * S.R), data=md, start=list(x=1, y=2, z=-3))
fit <- predict(mod, list(S.R=seq(0, 0.5, 0.01)))
lines(fit ~ seq(0, 0.5, 0.01))
# Get parameters of fit
formula(mod); coef(mod)
```

### PI curve
```{r run1_PI, cache=T, fig.height=3}
sym_PI(pars=defpars)
```

@@

# Run to steady state and plot results
ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kCO2"), c(1, 10)), at=at, food=0, runtime=T)
plot_steady_states(ssdat)
```
With default parameters, steady state values of specific growth are higher than usually observed in corals (skeletal growth), but patterns across environments are consistent with literature.

```{r ss_feeding, cache=T, fig.width=5, fig.height=4.4, eval=F, include=F}
# Set parameters
defpars <- def_pars()  # Get default parameters

@@

# Run to steady state and plot results
ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kCO2"), c(1, 10)), at=at, food=5e-6, runtime=T)
plot_steady_states(ssdat)
```
With default parameters, steady state values of specific growth are higher than usually observed in corals (skeletal growth), but patterns across environments are consistent with literature.

@@

# Run to steady state and plot results
ssdat <- run_steady_states(pars=replace(defpars, c("initS", "kROS", "kCO2"), c(1, 120, 10)), at=at, food=0, runtime=T)
plot_steady_states(ssdat)
```
With some feeding (more realistic), can maintain growth with very low symbionts (bleached) at very high or low light with high nutrients.

### Dynamic behavior: nutrients
```{r varyN, eval=F, include=F}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(20,20,2), N=c(1e-7,4e-6,2), X=c(0,0,0))

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.25, 10)))#
plot_run(run)
```

```{r varyN_lowL, eval=F, include=F}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

@@

### Dynamic behavior: light
```{r varyL, eval=T, include=T, fig.width=5}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
plot_run(run)
```

### Dynamic behavior: steadily increasing light
```{r incrL, eval=T, include=T, fig.width=5}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.1, 10)))#
plot_run(run)
```

### Dynamic behavior: steadily increasing light, + high N
```{r incrL_highN, eval=T, include=T, fig.width=5}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(30,60,0), N=c(2e-6,2e-6,2), X=c(5e-6,5e-6,0))

@@

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.15, 10)))#
plot_run(run)
```

### Dynamic behavior: high light
```{r varyHL, eval=T, include=T, fig.width=5}
# Set run time vector
time <- seq(1, 365, 0.5)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(5,45,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
plot_run(run)
```

### Dynamic behavior: high light - 3 years
```{r varyHL3, eval=T, include=T, fig.width=5}
# Set run time vector
time <- seq(1, 1095, 0.5)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(5,45,2), N=c(1e-7,1e-7,0), X=c(5e-6,5e-6,0))

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
plot_run(run)
```

### Dynamic behavior: high light - 3 years - no feeding
```{r varyHL3noF, eval=T, include=T, fig.width=5}
# Set run time vector
time <- seq(1, 1095, 0.5)  # if single model run, use time input

# Initialize environment
env <- init_env(time=time, L=c(20,45,2), N=c(1e-7,1e-7,0), X=c(0,0,0))

# Run simulation
run <- run_coral(time=time, env=env, pars=replace(defpars, c("initS", "kCO2"), c(0.2, 10)))#
plot_run(run)
```

### Dynamic behavior: high light & interacting factors
```{r incrL_multiline1, include=T, eval=T, fig.width=5}
time <- seq(1, 200, 0.5)  # Set time
defpars <- def_pars()  # Get default parameters
defpars <- replace(defpars, "kNPQ", 112)

# Run 1
env1 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,0), X=c(2e-6,2e-6,0))
run1 <- run_coral(time=time, env=env1, pars=replace(defpars, c("initS", "kCO2"), c(0.12, 8)))
# Run 2 - increase food
env2 <- init_env(time=time, L=c(15,40,2), N=c(1e-7,1e-7,0), X=c(8e-6,8e-6,0))
run2 <- run_coral(time=time, env=env2, pars=replace(defpars, c("initS", "kCO2"), c(0.12, 8)))
# Run 3 - increase DIN
env3 <- init_env(time=time, L=c(15,40,2), N=c(1e-6,1e-6,0), X=c(2e-6,2e-6,0))
run3 <- run_coral(time=time, env=env3, pars=replace(defpars, c("initS", "kCO2"), c(0.16, 8)))

# Plot
par(mfrow=c(1,1), mar=c(4,4,3,4))
plot(time, env1$L, col=alpha("gold", 0.5), type="l", lwd=2, axes=F, xlab="Time", ylab="S/H biomass ratio")
axis(side=4); mtext(side=4, text="Light (mol m-2 d-1)", line=2)
par(new=T)
plot(NA, xlim=range(time), ylim=c(0,0.2), yaxs="i", ylab="", xlab="")
with(run1, lines(time, S$S/H$H, lty=1))
with(run2, lines(time, S$S/H$H, lty=2))
with(run3, lines(time, S$S/H$H, lty=3))
legend("topright", legend=c("default", "+feeding", "+DIN"), lty=c(1,2,3))

@@
